---
title:  "Report"
author: "Luis Varela"
date:   "25 4 2020"
output: html_document
---


<div style="margin-bottom:100px;">
</div>

<h2 style="text-align: center;">Introduction</h2>

<p style="text-align: justify;">On this report, profitability of (a) certain article(s) will be demonstrated. The df for the task consists of 80 observations, from a df of existing products and 24 of a list of possible candidates.</p>

<div style="margin-bottom:100px;">
</div>

<h2 style="text-align: center;">Used Libraries</h2>


```{r results='asis', warning=FALSE, message=FALSE}
#libraries needed
library(arules)
library(arulesViz)
library(knitr)
library(ggplot2)
```


<div style="margin-bottom:100px;">
</div>

<h2 style="text-align: center;">Data</h2>


```{r results='markup', fig.align='center', out.extra='angle=90', message=FALSE}
#reading the transaction file. There was an error "incomplete final line found on 'Data/ElectronidexTransactions2017.csv" fixed by
#opening the csv file in the editor and pressing <enter>
transaction_list <- read.transactions("Data/ElectroIndexTrans2017.csv", 
                                      format = "basket", 
                                      sep=",", 
                                      rm.duplicates=TRUE)

#load product categories list
itemIndex <- read.csv("Data/ItemIndex.csv", sep=",")

#add level1 to categories our products
transaction_list@itemInfo$level1 <- itemIndex$ProductCategory

#summary of the df
summary(transaction_list)

#find item that was consumed alone
oneItem <- transaction_list[which(size(transaction_list) == 1), ] 
oneItem

#remove "" from the transaction list to avoid it giving us wrong results 
transaction_list@itemInfo$labels <- gsub("\"", "", transaction_list@itemInfo$labels)
```


## Observations

* the density was 0.035 means 3.5% are non zero matrix cells

* the matrix has 9835 times 125, i.e. 1229375 cells. So, if we want to know how many items were bought, we have 9835 times 125 times 0.03506172 it gives us that 43104 items were bought

* Average transaction contained 43104/9835 = 4,38 items

* The first quartile and median purchase size are 2 and 3 items respectively, implying that 25 percent of transactions contained two or fewer items and about half contained around three items

* 2163 items were consumed alone


<div style="margin-bottom:100px;">
</div>

<h2 style="text-align: center;">Analysis</h2>


```{r message=FALSE, fig.align='center'}
#lists items on transaction files
itemLabels(transaction_list) # We have 249 different items

#item frequency with a 0.1 support
itemFrequencyPlot(transaction_list, 
                  support=0.1)

#item frequency with a 0.5 support
itemFrequencyPlot(transaction_list, 
                  support=0.05)

#plot frequency of each item (absolute)
itemFrequencyPlot(transaction_list, 
                  topN=10, 
                  main='Absolute Item Frequency Plot', 
                  type="absolute", 
                  ylab="Item Frequency (Absolute)")

#visualizing random 100 transactions
image(sample(transaction_list, 100))
```


```{r, fig.align='center', out.extra='angle=90', message=FALSE}
#using the Apriori alg. with support at 0.01 and confidence at 0.4. 
#There will be other tryals to compare how changing one or the other affects the results
purchaseRules <- apriori(transaction_list,
                         parameter = list(sup = 0.01,
                                          conf = 0.4,
                                          minlen = 2,
                                          target = "rules"))

#inspect resulting rules
inspect(head(purchaseRules, n = 10))

#taking a closer look
inspect(sort(purchaseRules, by = "lift", decreasing = TRUE)[1:10]) #interesting show of the association with HP Laptops

#sorting by sonfidence
inspect(sort(purchaseRules, by = "confidence", decreasing = TRUE)[1:10]) #interesting show of the association with HP Laptops

#simple plot of the rules
plot(purchaseRules, jitter = 0) #confidence remains high when support is low, this should be kept in mind

#ploting rules
plot(purchaseRules, col = c("red", "blue"), shading = "order", control = list(main = "Two-key plot"), jitter = 0)
```


```{r results='markup', fig.align='center', out.extra='angle=90', message=FALSE}
#checking how support and confidence affect the effectiveness of the rules
differentConf <- read.csv("Data/differentConfidence.csv", sep=",")
differentsupport <- read.csv("Data/differentSupport.csv", sep=",")

#correlation between number of rules and support level
ggplot(data = differentsupport, aes(x=min_support, y=number_rules)) + #the greater the number of rules the smaller the support
      geom_point() +
      geom_smooth(se=FALSE, color="Red") +
      xlab("Mean Support") + 
      ylab("Number of Rules") +
      ggtitle("Mean Support against Rules found (confidence = 0.4)") +
      theme_light()

#correlation between number of rules and confidence level
ggplot(data = differentConf, aes(x=minimum_confidence, y=number_rules)) + #the more rules we have the least confidence we would get
      geom_point() +
      geom_smooth(se=FALSE, color="Red") +
      xlab("Mean Confidence") + 
      ylab("Number of Rules") +
      ggtitle("Mean Confidence against Rules found (support = 0.01)") +
      theme_light()
```


```{r, set-options, message=FALSE, fig.align='center', cache=FALSE}
options(witdh = 2600)
#here is where there will be a bridge between this excercise and the last one
#I found that only 3 product types make the cut generating 336 rules
tempDesktop <- c("iMac", "Dell Desktop", "Lenovo Desktop Computer")
tempLaptop <- c("HP Laptop")
tempMonitors <- c("ViewSonic Monitor")

#running Apriori to find out how many rules we have
rulesAll <- apriori(transaction_list, parameter = list(sup = 0.005, 
                                          conf = 0.4, 
                                          minlen = 2, 
                                          maxlen = 20))

#subsetting rules with a where people are more likely to close their purchase + a Desktop
ruleDesktop <- subset(rulesAll, subset = rhs %in% tempDesktop)

#removing redundant rules
ruleDesktop <- ruleDesktop[!is.redundant(ruleDesktop)] #10 rules were found to be redundat
summary(ruleDesktop)

#top 15 support/conf/lift
inspect(sort(ruleDesktop, decreasing = TRUE, by = "support")[1:15])
inspect(sort(ruleDesktop, decreasing = TRUE, by = "confidence")[1:15])
inspect(sort(ruleDesktop, decreasing = TRUE, by = "lift")[1:15])

#subsetting rules with a where people are more likely to close their purchase + a Laptop
ruleLaptop <- subset(rulesAll, subset = rhs %in% tempLaptop)

#remove redundant rules
ruleLaptop <- ruleLaptop[!is.redundant(ruleLaptop)] #2 ruöes were found to be redundant
summary(ruleLaptop)

inspect(sort(ruleLaptop, decreasing = TRUE, by = "support")[1:15])
inspect(sort(ruleLaptop, decreasing = TRUE, by = "confidence")[1:15])
inspect(sort(ruleLaptop, decreasing = TRUE, by = "lift")[1:15])

#subsetting rules with a where people are more likely to close their purchase + a Monitor
ruleMonitors <- subset(rulesAll, subset = rhs %in% tempMonitors)

#remove redundant rules
ruleMonitors <- ruleMonitors[!is.redundant(ruleMonitors)] #all rules remained
summary(ruleMonitors)

#Sort by top 15 support/conf/lift
inspect(sort(ruleMonitors, decreasing = TRUE, by = "support"))
inspect(sort(ruleMonitors, decreasing = TRUE, by = "confidence"))
inspect(sort(ruleMonitors, decreasing = TRUE, by = "lift"))

#aggregate by cats
transaction_list_byType <- aggregate(transaction_list, by = transaction_list@itemInfo$level1)

#plot items frequency for categories
itemFrequencyPlot(transaction_list_byType,
                  topN = 10,
                  main = 'Absolute Item Frequency Plot',
                  type = "absolute",
                  ylab = "Item Frequency (Absolute)")

ruleByType <- apriori(transaction_list_byType, 
                      parameter = list(sup = 0.005, 
                                       conf = 0.4, 
                                       minlen = 3, 
                                       maxlen = 20))

ruleByBWCats <- subset(ruleByType, subset = rhs %in% c("Desktop", 
                                                       "Laptops", 
                                                       "Monitors",
                                                       "Accessories", 
                                                       "Computer Tablets",
                                                       "Printers"))

ruleByBWCats <- subset(ruleByType, subset = rhs %in% c("Desktop") & lift > 1.5)

#remove duplicates
ruleByBWCats <- ruleByBWCats[!is.redundant(ruleByBWCats)] #5 duplicated were found
summary(ruleByBWCats)

#Sort by top 15 support/conf/lift to explore
inspect(sort(ruleByBWCats, decreasing = TRUE, by = "support")[1:15])
inspect(sort(ruleByBWCats, decreasing = TRUE, by = "confidence")[1:15])
inspect(sort(ruleByBWCats, decreasing = TRUE, by = "lift")[1:15])


#note to self: bad support and medium confidence, only lift is good. Will be very rare for the rules to happen
```


```{r}
#for reproducibility
sessionInfo()
```